### 2019-11-14 fragments/yang-ma-li.tex
cplx-transformer review

Speech recognition, signal processing, audio transcription all use FFT to get
spectral input, which is complex.

cplx-nets have rich represeantaiton

(lit-review)
attn and xfromer gained prominence in seq. modelling especially with long
temporal dependencies [10-13]

cplx-feed forward and cplx-lstm (with unitary matrix constraints) [see. 3-10]
cplx-conv and BN [18]
cplx-recurrent [5-7,9,10] merging FT and seq modelling [8]

(cplx)
natural representation of complex numbers as real pairs
emulate complex algebra and operations using paired reals

(attn)
Q, K, V = X W_Q, X W_K, X W_V
A = Q K^T V

use min-max attn instead of softmax (grad explosion !!)
are matrices W cplx?

% datasets
MusicNet : audio (signal) to symbols (labels), resampling 44khz to 11khz according to
[18], then real signal to cplx via FT

in-phase quadratic (private dataset): classify device by WIFI signal, resample to
balance devices.