### 2019-11-14 fragments/monning-manandhar.tex
capacity comparison of cplx nets summary

the paper studies capacity of simple complex networks
conclude that cplx nets perform not better than real-valued netowrks with similar capacity.
Observe synchronisation of real and imaginary parts in real valued tasks, but observe subpar
performance of cplx netowrks of equal capacity, sensetivity ot initialization. Isues due to 
singularities (poles) -- avoidable by constraining function value, normalizing weights or
gradients. Found that using relus is stabler and allows insight transfer from the vast literature
on real nets to complex domain. Recommend using cplx only if the data naturally or meaningfully
maps into cplx plane. and the net should reflect real-imag interaction

ensure comparability of architectures with respect to model size and capacity: doubling of
the number of parameters implies that one should keep the number of numerical parameters as
close as possible whn comparing real and complex nets. This ensures that performance differences
are due to use of cplx numbers rather than doubling of the number of parameters.
1) fixed number of real-valued parameters per layer
2) fixed budget of real-valued parameters per network

(lit-review) formally described in [8], complex versions of backpropagation in [6, 10, 19],
ides were even extended to quaternions [4, 3], complex valued RNN [5, 29] with unitary matrices
to tackle vanishing/exploding gradient problems. Complex convolutions [26], and complex attention
layers for complex-valeud transformer achitectures [ynag-ma-li]
Complex version of  batch-normalisation and weight initialization [26] (analogue of glorot xavier)

cplx-nets had limited popularity: less intuitive architecutres, increased numerical complexity (in
paired real -- instead of a single up to four multiplications are required, and two additions).
% due to them not being true cplx and just paired real computational graphs

(tasks) processing and analysis of compelx-valued data or data with intuitive mapping to
complex numbers. Images and signals in spectral representation. [15] show superiority of complex
nets to such tasks.

(datasets) MNIST; synthetic-1 (quadrant of the sum of cplx); synthetic-2 (same af synth-1 but for real)

(cplx) activation function can be C->C or C->R. cplx-differentiable -- exists limit in
C-field -- holomorhpic. Wirtinger CR calculus -- chosing a different basis of partial derivatives
CR improves statbility of the training process.

Activations: separate application to real and imaginary part, or more sophisticated, which uses
interaction between cplx and its conjugate, be it polar or paired representation. modReLU, hyperbolic
functions, complex modulus (absolute value), softmax based on squared modulus

simple doubling in insufficient: we also need to duplicate the parameters and enforce skew symmetric
constraint ([[re, -im], [im, re]])

complex-valued loss functions need a notion of ordering (what does it mean to minimize it,
how is it related to ERM and stat learning)

[03] Aizenberg (2007) Multilayer feedfroward network on multi-valued neurons and back prop
[04] Aizenberg (1992) Cnn based on muli-valeud neurons 
[05] Arjovsky, Shah (2015) Unitary evolution recurrent neural networks
[06] Benvenuto, Piazza (1992) complex backprop algorithm
[08] Clarke (1990) Generalization on neural networks to compelx plane
[10] Gergious Koutsougeras (1992) Complex domain backpropagation
[15] Hirose (2009) Cmoplex valued neural networks
[19] Nitta (1993) A back-propagation algorithm for compelx numbered neural networks
[26] Trabelsi (2017) Deep complex networks
[29] Wisdom Powers (2016) Full-capacity unitary recurrent neural networks