### 2019-11-14 fragments/yang-ma-li.tex

Speech recognition, signal processing, audio transcription all use FFT to get
spectral input, which is complex.

cplx-nets have rich represeantaiton

(lit-review)
attn and xfromer gained prominence in seq. modelling especially with long
temporal dependencies [10-13]

cplx-feed forward and cplx-lstm (with unitary matrix constraints) [see. 3-10]
cplx-conv and BN [18]
cplx-recurrent [5-7,9,10] merging FT and seq modelling [8]

(cplx)
natural representation of complex numbers as real pairs
emulate complex algebra and operations using paired reals

(attn)
Q, K, V = X W_Q, X W_K, X W_V
A = Q K^T V

use min-max attn instead of softmax (grad explosion !!)
are matrices W cplx?

% datasets
MusicNet : audio (signal) to symbols (labels), resampling 44khz to 11khz according to
[18], then real signal to cplx via FT

in-phase quadratic (private dataset): classify device by WIFI signal, resample to
balance devices.

[03] Hirose (2003) Cmoplex valued neural networks: theories and applications
[04] Amin, Murase (2009) single layered complex-valued neural networks for real-valued classification problems
[05] Arjovsky, Shah (2016) Unitary evolution recurrent neural networks
[06] Danihelka, Wayne (2016) Associative Long short-term memory
[07] Wolter, Yao (2018) Complex gatteed recurrent nural network
[08] Wolter, Yao (2018) Fourier RNNs for sequence prediction
[09] Hochreiter, Schmidhuber (1997) Long short-term memory
[10] Cho, Merrienboer (2014) Learning phase representations using RNN encode-decoder for statistical machine translation
[11] Bahndau, Cho (2014) Neural Machine translation by jointly learning to align and translate
[12] Vaswani, Shazeer (2017) Attention is all you need
[13] Yang, Dai (2019) Xlnet: generalized autoregressive pretraining for language understsding
[18] Trabelsi (2017) Deep complex networks