### 2020-01-13 fragments/wisdom_et_al_2016.tex
minor review of Wisdom-etal2016

(wisdom_full-capacity_2016)
Recurrent nn suffer from vanishing/exploding gradients due to repeated use of the parameter matrix
which affects forward pass and backprop via layers' input gradients.

remedies included gradient clipping, orthogonal initializations, novel architectures with
gated, residual, or attention based pathways.

unitary recurrent weight matrices make the gradient inherently stable \citep{arjovsky_unitary_2016}
Illustrate on pixel-by-pixel mnist

This paper shows the limitations of Arjovsky's parameterization and argues in favour of fully fledged
optimization on the Stiefel manifold, i.e. the differentiable manifold of $\cplx^{n \times n}$ unitary
matrices.