### 2020-01-13 fragments/wisdom_et_al_2016.tex
merged in fragments of arjovsky et al 2016

(wisdom_full-capacity_2016)
Recurrent nn suffer from vanishing/exploding gradients due to repeated use of the parameter matrix
which affects forward pass and backprop via layers' input gradients.

remedies included gradient clipping, orthogonal initializations, novel architectures with
gated, residual, or attention based pathways.

unitary recurrent weight matrices make the gradient inherently stable \citep{arjovsky_unitary_2016}
Illustrate on pixel-by-pixel mnist

This paper shows the limitations of Arjovsky's parameterization and argues in favour of fully fledged
optimization on the Stiefel manifold, i.e. the differentiable manifold of $\cplx^{n \times n}$ unitary
matrices.

(arjovsky_unitary_2016)
vanishing gradients adversely affect credit assignment function of backpropagation
explicitly derive the upper bound on the norm of the gradient wrt intermediate recurrent outputs
as a a function of depth (through extensive use of the chain-rule, and the definition of the
operator norm of a square matrix; and if intermediate output grads die/explode, then the grad wrt
the weight of the layers tends to die/explode). Thus they motivate the use of unitary (orthogonal)
weight matrices.

Note that in their experiments, split-re-im ReLU adversely affected the performance by
``brutally'' destroying the pahse, and so the decision was made to use modReLU, that applies
the rectifier to the amplitude of a complex number.

Finally, they use a mixing layer to transform complex to real outputs as in \cite{wolter_yao_2018}.